---
title: "Week 7 - Functions"
author: "Angie Bouche, Joslyn Fritz"
date: "May 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Average power for cars under normal conditions

```{r sampling}

source("autopower.R") #take function from R file

# generate sample speeds from a distribution with 10% sd
nsample = 100
speeds = rnorm(mean=28, sd=2.8, n=nsample)

#generate a data frame from the model
results = data.frame(speed=speeds, power=NA)

head(results)

results$power = autopower(V=speeds, A=25, m=20000)

#Use graphs to check
graph_1<- ggplot(results, aes(speed, power/1000))+geom_point()+labs(x="speed in m/s", y="power in kW")
graph_2 <-ggplot(results, aes(x="Across Speed",y=power/1000))+geom_boxplot()+labs(y="power in kW")
graph_1
graph_2


#select mass and area of cars
possible_cars = data.frame(mass=c(31000,45000,38000), area = c(25,30,22))
 

#enter base speed at 100 kmh and convert to m/s
speed_base = 100 * 0.28

#set power equation for three cars selected above
possible_cars$power = autopower(V=speed_base, A = possible_cars$area, m=possible_cars$mass)
 

possible_cars$prob = c(0.4, 0.4, 0.2)

#assume normal distribution and set variable for speed assuming 100 kmh with 10% sd
mn = log(100*0.277)
nsample = 100
speeds = rlnorm(mean=mn, sd=0.1*mn, nsample)
summary(speeds)
 

results = data.frame(speed=speeds, power=NA)


#select all possible cars in distribution
possible_cars$row = seq(from=1, to=nrow(possible_cars))
possible_cars

 
whichcar = base::sample(possible_cars$row, size=nsample, prob=possible_cars$prob, replace=TRUE)
whichcar #shows you 100 observations choosing a number between 1 and 3 since three car types are listed

#this is basically bootstrapping


head(whichcar)#begining of list of possible car selections


#show masses for selected cars in distribution
results$mass = possible_cars$mass[whichcar]

head(results)

# Add areas for selected cars
results$area = possible_cars$area[whichcar]

 

# power for all of our samples of speed and car type

results$power = autopower(A=results$area, V=results$speed, m=results$mass)

summary(results$power)
ggplot(results,aes(x="", y=power/1000))+geom_boxplot(fill="red")+labs("Power kW")



```

## Average power for cars when rolling coefficient is reduced by half
```{r autopower_cr}
source("autopower_cr.R")

# generate sample speeds from a distribution with 10% sd
nsample_cr = 100
speeds_cr = rnorm(mean=28, sd=2.8, n=nsample_cr)

#generate a data frame from the model
results_cr = data.frame(speed=speeds_cr, power=NA)

head(results_cr)

results_cr$power = autopower(V=speeds, A=25, m=20000)

#Use graphs to check
graph_1_cr<- ggplot(results_cr, aes(speed, power/1000))+geom_point()+labs(x="speed in m/s", y="power in kW")
graph_2_cr <-ggplot(results_cr, aes(x="Across Speed",y=power/1000))+geom_boxplot()+labs(y="power in kW")
graph_1_cr
graph_2_cr


#select mass and area of cars
possible_cars_cr = data.frame(mass=c(31000,45000,38000), area = c(25,30,22))
 

#enter base speed at 100 kmh and convert to m/s
speed_base_cr = 100 * 0.28

#set power equation for three cars selected above
possible_cars_cr$power = autopower(V=speed_base_cr, A = possible_cars_cr$area, m=possible_cars_cr$mass)
 

possible_cars_cr$prob = c(0.4, 0.4, 0.2)

#assume normal distribution and set variable for speed assuming 100 kmh with 10% sd
mn_cr = log(100*0.277)
nsample_cr = 100
speeds_cr = rlnorm(mean=mn, sd=0.1*mn, nsample)
summary(speeds_cr)
 

results_cr = data.frame(speed=speeds_cr, power=NA)


#select all possible cars in distribution
possible_cars_cr$row = seq(from=1, to=nrow(possible_cars_cr))
possible_cars_cr

 
whichcar_cr = base::sample(possible_cars_cr$row, size=nsample_cr, prob=possible_cars_cr$prob, replace=TRUE)
whichcar_cr #shows you 100 observations choosing a number between 1 and 3 since three car types are listed

#this is basically bootstrapping


head(whichcar_cr)#begining of list of possible car selections


#show masses for selected cars in distribution
results_cr$mass = possible_cars_cr$mass[whichcar_cr]

head(results_cr)

# Add areas for selected cars
results_cr$area = possible_cars_cr$area[whichcar_cr]

 

# power for all of our samples of speed and car type

results_cr$power = autopower(A=results_cr$area, V=results_cr$speed, m=results_cr$mass)

summary(results_cr$power)
ggplot(results_cr,aes(x="", y=power/1000))+geom_boxplot(fill="red")+labs("Power kW")




```

##Average power for cars when speed is reduced to 80 kmh on average (Assuming 10% standard deviation)
```{r autpower_sp}
source("autopower.R") #take function from R file

# generate sample speeds from a distribution with 10% sd
nsample_sp = 100
speeds_sp = rnorm(mean=22.4, sd=2.24, n=nsample_sp)

#generate a data frame from the model
results_sp = data.frame(speed=speeds_sp, power=NA)

head(results_sp)

results_sp$power = autopower(V=speeds_sp, A=25, m=20000)

#Use graphs to check
graph_1<- ggplot(results_sp, aes(speed, power/1000))+geom_point()+labs(x="speed in m/s", y="power in kW")
graph_2 <-ggplot(results_sp, aes(x="Across Speed",y=power/1000))+geom_boxplot()+labs(y="power in kW")
graph_1
graph_2


#select mass and area of cars
possible_cars_sp = data.frame(mass=c(31000,45000,38000), area = c(25,30,22))
 

#enter base speed at 80 kmh and convert to m/s
speed_base_sp = 80 * 0.28

#set power equation for three cars selected above
possible_cars_sp$power = autopower(V=speed_base_sp, A = possible_cars_sp$area, m=possible_cars_sp$mass)
 

possible_cars_sp$prob = c(0.4, 0.4, 0.2)

#assume normal distribution and set variable for speed assuming 80 kmh with 10% sd
mn_sp = log(80*0.277)
nsample_sp = 100
speeds_sp = rlnorm(mean=mn_sp, sd=0.1*mn_sp, nsample_sp)
summary(speeds_sp)
 

results_sp = data.frame(speed=speeds_sp, power=NA)


#select all possible cars in distribution
possible_cars_sp$row = seq(from=1, to=nrow(possible_cars_sp))
possible_cars_sp

 
whichcar_sp = base::sample(possible_cars_sp$row, size=nsample_sp, prob=possible_cars_sp$prob, replace=TRUE)
whichcar_sp #shows you 100 observations choosing a number between 1 and 3 since three car types are listed

#this is basically bootstrapping


head(whichcar_sp)#begining of list of possible car selections


#show masses for selected cars in distribution
results_sp$mass = possible_cars_sp$mass[whichcar_sp]

head(results_sp)

# Add areas for selected cars
results_sp$area = possible_cars_sp$area[whichcar_sp]

 

# power for all of our samples of speed and car type

results_sp$power = autopower(A=results_sp$area, V=results_sp$speed, m=results_sp$mass)

summary(results_sp$power)
ggplot(results_sp,aes(x="", y=power/1000))+geom_boxplot(fill="red")+labs("Power kW")




```


## Assignment 3:

Based on our analysis, the mean power used over a normal distribution of car weight and surface areas, assuming a speed limit of 100km/h with standard deviation of 10%, rolling coefficient due to friction of 0.015, coefficient due to drag of 0.3, and air pressure of 1.2 kg/m^3 is `r round(mean(results$power), 0)`w. If we reduce the rolling coefficient/friction of the road by half (to 0.0075) the mean power is reduced to `r round(mean(results_cr$power),0)`w. Alternatively, if we reduce the speed limit of the road to 80 km/h the mean power is reduced to `r round(mean(results_sp$power),0)`w. Lowering the speed limit is more effective at reducing power needed, and this technique would be less expensive to implement. 