---
title: "Assignment 2 RMD"
author: "Angie Bouche"
date: "May 17, 2019"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse) #load tidyverse
library(DBI)

```

#Part 1: Import and Tidy

#1. Load raw data into `parcels_raw` tibble
```{r}
parcels_raw <-   read_delim(
    "Santa_Barbara_County_parcels_2011.csv",
    delim = ",",
    col_types = cols(.default = col_character()) #read in parcels data csv and convert to tibble
  )

parcels_raw
```

#2/3. Create `parcels` tibble with only certain columns. Change na values. Convert columns to appropriate type
```{r}

parcels<- parcels_raw %>% 
  select(UseCode, APN, Situs1, Situs2, Acreage, NonTaxCode, AgPres, LandValue, Net_Impr, Net_AV, M_Address1, M_Address2) %>% #select appropriate columns for parcels tibble
  mutate(Net_AV, ifelse(Net_AV<0, NA, Net_AV), na = "", na = "0")#remove any negative assessment values, change 'space' and '0' placeholders to NA
  

parcels$Acreage=parse_double(parcels$Acreage) #convert numerical data columns from characters
parcels$LandValue= parse_double(parcels$LandValue)
parcels$Net_Impr = parse_double(parcels$Net_Impr)
parcels$Net_AV = parse_double(parcels$Net_AV)


parcels

```

#4. Format data into an SQL Database file
```{r}
parcelsdb <- dbConnect(RSQLite::SQLite(), dbname = "parcels.db")#connect Rmd SQL chunk to SQLite database

dbWriteTable(conn = parcelsdb, name = "parcels", value = parcels, overwrite = TRUE) # Turn tibble into a database. first argument is conenction handle, second is name for table in database, third is the dataframe itself. Overwrite to allow the code to be run multiple times.

```

#Part 2: Analyze

```{r}

usecodes<-read_delim(
    "USeCodes.csv",
    delim = "|",
    col_types = cols(.default = col_character()) #read in parcels data csv and convert to tibble
  ) %>% 
  rename (UseCode2 = UseCode)

dbWriteTable(conn = parcelsdb, name = "usecodes", value = usecodes, overwrite = TRUE) # TUrn tibble into a database. first argument is conenction handle, second is name for table in database, third is the dataframe itself
```


#1. 10 most frequently occurring land uses (descending order)
```{sql connection=parcelsdb}

SELECT UseCode, CodeDesc, count(UseCode) AS "Number of Properties"
FROM   parcels
JOIN usecodes ON Parcels.UseCode = usecodes.UseCode2
GROUP BY UseCode
ORDER By count(UseCode) DESC
LIMIT 10;

```

#2. How many acres are in agricultural preserves?

```{sql connection=parcelsdb}

SELECT round(sum(Acreage),2) AS "Total Acreage" FROM parcels WHERE AgPres >0; 

```
Only count cells where AgPres is over 0 to get rid of non-agricultural properties.

#3. What is the mean net assessed value per acre of the entire county?

```{sql connection=parcelsdb}
SELECT round(sum(Net_AV)/ sum(Acreage), 2) AS "Average Assessed Value per Acre" FROM parcels;

```


#4. What is the total net assessed value of all non-taxable parcels?

```{sql connection=parcelsdb}
SELECT sum(Net_AV) AS "Total Assessed Value" FROM parcels WHERE NonTaxCode >0;
```
Only count cells where NonTaxCode is over 0 to get rid of taxed properties.

#5.What are the 10 largest property holders, by acreage?
```{sql connection=parcelsdb}

SELECT M_Address1, M_Address2, sum(Acreage) AS "Total Acreage of Properties"
FROM parcels
WHERE M_Address1 IS NOT NULL AND M_Address2 IS NOT NULL 
GROUP BY M_Address1, M_Address2
ORDER BY  "Total Acreage of Properties" DESC
LIMIT 10; 
```
Only count cells where M_Address1 and M_Address2 is not null to get rid of properties with no associated address.

#6. What are the 10 largest property holders, by net assessed value?

```{sql connection=parcelsdb}

SELECT M_Address1, M_Address2,sum(Net_AV) AS "Total Assessed Value of Properties"
FROM parcels
WHERE  M_Address1 IS NOT NULL AND M_Address2 IS NOT NULL 
GROUP BY M_Address1, M_Address2
ORDER BY  "Total Assessed Value of Properties" DESC
LIMIT 10;
```

Only count cells where M_Address1 and M_Address2 is not null to get rid of properties with no associated address.