---
title: "Assignment 2 RMD"
author: "Angie Bouche"
date: "May 14, 2019"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse) #load tidyverse
library(DBI)


```

#Part 1: Import and Tidy

#1. Load raw data into `parcels_raw` tibble
```{r}
parcels_raw <-   read_delim(
    "Santa_Barbara_County_parcels_2011.csv",
    delim = ",",
    col_types = cols(.default = col_character()) #read in parcels data csv and convert to tibble
  )

parcels_raw
```

#2/3. Create `parcels` tibble with only certain columns. Change na values. Convert columns to appropriate type
```{r}

parcels<- parcels_raw %>% 
  select(UseCode, APN, Situs1, Situs2, Acreage, NonTaxCode, AgPres, LandValue, Net_Impr, Net_AV, M_Address1, M_Address2) %>% #select appropriate columns for parcels tibble
  mutate(Net_AV, ifelse(Net_AV<0, NA, Net_AV), na = "", na = "0")#remove any negative assessment values, change 'space' and '0' placeholders to NA
  

parcels$Acreage=parse_double(parcels$Acreage) #convert numerical data columns from characters
parcels$LandValue= parse_double(parcels$LandValue)
parcels$Net_Impr = parse_double(parcels$Net_Impr)
parcels$Net_AV = parse_double(parcels$Net_AV)


parcels

```

#4. Format data into an SQL Database file
```{r}
parcelsdb <- dbConnect(RSQLite::SQLite(), dbname = "parcels.db")#connect Rmd SQL chunk to SQLite database

dbWriteTable(conn = parcelsdb, name = "parcels", value = parcels) # first argument is conenction handle, second is name for table in database, third is the dataframe itself

dbListTables(parcelsdb)#check table
dbDisconnect(parcelsdb)#close


```

#Part 2: Analyze

#1. 10 most frequently occurring land uses (descending order)
```{sql connection=db}


SELECT *
FROM   parcels JOIN use_codes
ON     UseCode = UseCode;

SELECT UseCOde FROM Parcels ORDER BY quant DESC;

```

#2. How many acres are in agricultural preserves?

```{sql connection=db}

SELECT sum(Acreage) FROM parcels WHERE AgPres >0


```

#3. What is the mean net assessed value per acre of the entire county?

```{sql connection=db}
SELECT sum(Net_AV)/ sum(Acreage) FROM parcels

```


#4. What is the total net assessed value of all non-taxable parcels?

```{sql connection=db}
SELECT sum(Net_AV) FROM parcels WHERE NonTaxCode >0
```


#5.What are the 10 largest property holders, by acreage?
```{r}
topten_acres <- parcels %>% 
  filter(M_Address1 != "NA", M_Address2 != "NA") %>% #filter to include only properties with an owner
  group_by(M_Address1, M_Address2) %>% #group by owner
  summarize(sum_col = sum(Acreage)) %>%  
  arrange(-sum_col) %>% 
  head(10) #add up acreage for each owner, arrange in descending order, select top 10 owners and their land mass
  
topten_acres

```

```{sql connection=db}

SELECT sum(Acreage) FROM parcels
WHERE M_Address1 AND M_Address2 != 'NA' 
GROUP BY M_Address1
ORDER BY  Acreage DESC
```


#6. What are the 10 largest property holders, by net assessed value?
```{r}

topten_acres <- parcels %>% 
  select(Net_AV, M_Address1, M_Address2) %>% 
  filter(M_Address1 != "NA", M_Address2 != "NA") %>% #filter to include only properties with an owner
  group_by(M_Address1, M_Address2) %>% #group by owner
  summarize(sum_col2 = sum(Net_AV)) %>%  
  arrange(-sum_col2) %>%
  head(10) #sum up value of property for each owner, arrange by descending order, select top 10 owners
  
topten_acres

```

```{sql connection=db}

SELECT sum(Net_AV) FROM parcels
WHERE M_Address1 != NULL AND
WHERE M_Address2 != NULL
ORDER BY Net_AV DESC
```
