---
title: "ESM 263 Assignment 1 RMD"
author: "Angie Bouche"
date: "May 7, 2019"
output:
  html_document: default
  pdf_document: default
---

#Part 1: Import and Tidy
```{r setup, include=FALSE}
library(tidyverse) #load tidyverse

```

#1. Load raw data into `parcels_raw` tibble
```{r}
parcels_raw <-   read_delim(
    "Santa_Barbara_County_parcels_2011.csv",
    delim = ",",
    col_types = cols(.default = col_character()) #read in parcels data csv and convert to tibble
  )

parcels_raw
```

#2/3. Create `parcels` tibble with only certain columns. Change na values. Convert columns to appropriate type
```{r}

parcels<- parcels_raw %>% 
  select(UseCode, APN, Situs1, Situs2, Acreage, NonTaxCode, AgPres, LandValue, Net_Impr, Net_AV, M_Address1, M_Address2) %>% #select appropriate columns for parcels tibble
  mutate(Net_AV, ifelse(Net_AV<0, NA, Net_AV), na = "", na = "0")#remove any negative assessment values, change 'space' and '0' placeholders to NA
  

parcels$Acreage=parse_double(parcels$Acreage) #convert numerical data columns from characters
parcels$LandValue= parse_double(parcels$LandValue)
parcels$Net_Impr = parse_double(parcels$Net_Impr)
parcels$Net_AV = parse_double(parcels$Net_AV)


parcels

```
#4. Format data back into csv from the tibble using "|" as a deliminator
```{r}
wrote <- write_delim(parcels,
  "final_parcels.csv",
  delim = "|",
  na = "") #create finalized csv using | deliminator

wrote

```

#Part 2: Analyze

#1. 10 most frequently occurring land uses (descending order)
```{r}

use_codes <-read_delim(
    "UseCodes.csv",
    delim = "|",
    col_types = cols(.default = col_character()) 
  ) #read in use codes data as tibble


parcel_use_code <- parcels %>% #merge use code and parcel datasetes
 left_join(use_codes, by = "UseCode") 

land_use <- parcel_use_code %>% 
  count(CodeDesc) %>% 
  arrange(-n) %>% 
 head(10) #count types of land use codes, arrange in descending order, select top 10 rows

land_use

```

#2. How many acres are in agricultural preserves?
```{r}

ag<- parcels %>% 
  filter(AgPres!= "NA") %>% 
  summarize(sum(Acreage)) #select agricultural preserve rows, and sum

ag

```

#3. What is the mean net assessed value per acre of the entire county?
```{r}
mean_nav<-parcels %>% 
  select(Acreage, Net_AV) %>% 
  summarize (mean = (sum(Net_AV)/sum(Acreage))) #select acreage and net assessed value columns, and average assessed value per acre
  
mean_nav
  
```

#4. What is the total net assessed value of all non-taxable parcels?
```{r}
nav_nontax <- parcels %>% 
  select(NonTaxCode, Net_AV) %>%
  filter(NonTaxCode != "NA", Net_AV != "NA") %>% 
  summarize(sum(Net_AV))

nav_nontax

```

#5.What are the 10 largest property holders, by acreage?
```{r}
topten_acres <- parcels %>% 
  filter(M_Address1 != "NA", M_Address2 != "NA") %>% #filter to include only properties with an owner
  group_by(M_Address1, M_Address2) %>% #group by owner
  summarize(sum_col = sum(Acreage)) %>%  
  arrange(-sum_col) %>% 
  head(10) #add up acreage for each owner, arrange in descending order, select top 10 owners and their land mass
  
topten_acres

```

#6. What are the 10 largest property holders, by net assessed value?
```{r}

topten_acres <- parcels %>% 
  select(Net_AV, M_Address1, M_Address2) %>% 
  filter(M_Address1 != "NA", M_Address2 != "NA") %>% #filter to include only properties with an owner
  group_by(M_Address1, M_Address2) %>% #group by owner
  summarize(sum_col2 = sum(Net_AV)) %>%  
  arrange(-sum_col2) %>%
  head(10) #sum up value of property for each owner, arrange by descending order, select top 10 owners
  
topten_acres

```

