---
title: "Assignment 2 RMD"
author: "Angie Bouche"
date: "May 14, 2019"
output: html_document
---


```{r setup, include=FALSE}
library(tidyverse) #load tidyverse
library(DBI)

```

#Part 1: Import and Tidy

#1. Load raw data into `parcels_raw` tibble
```{r}
parcels_raw <-   read_delim(
    "Santa_Barbara_County_parcels_2011.csv",
    delim = ",",
    col_types = cols(.default = col_character()) #read in parcels data csv and convert to tibble
  )

parcels_raw
```

#2/3. Create `parcels` tibble with only certain columns. Change na values. Convert columns to appropriate type
```{r}

parcels<- parcels_raw %>% 
  select(UseCode, APN, Situs1, Situs2, Acreage, NonTaxCode, AgPres, LandValue, Net_Impr, Net_AV, M_Address1, M_Address2) %>% #select appropriate columns for parcels tibble
  mutate(Net_AV, ifelse(Net_AV<0, NA, Net_AV), na = "", na = "0")#remove any negative assessment values, change 'space' and '0' placeholders to NA
  

parcels$Acreage=parse_double(parcels$Acreage) #convert numerical data columns from characters
parcels$LandValue= parse_double(parcels$LandValue)
parcels$Net_Impr = parse_double(parcels$Net_Impr)
parcels$Net_AV = parse_double(parcels$Net_AV)


parcels

```

#4. Format data into an SQL Database file
```{r}
parcelsdb <- dbConnect(RSQLite::SQLite(), dbname = "parcels.db")#connect Rmd SQL chunk to SQLite database

dbWriteTable(conn = parcelsdb, name = "parcels", value = parcels, overwrite = TRUE) # first argument is conenction handle, second is name for table in database, third is the dataframe itself

```

#Part 2: Analyze

```{r}

usecodes<-read_delim(
    "USeCodes.csv",
    delim = "|",
    col_types = cols(.default = col_character()) #read in parcels data csv and convert to tibble
  ) %>% 
  rename (UseCode2 = UseCode)

dbWriteTable(conn = parcelsdb, name = "usecodes", value = usecodes, overwrite = TRUE) # first argument is conenction handle, second is name for table in database, third is the dataframe itself
```


#1. 10 most frequently occurring land uses (descending order)
```{sql connection=parcelsdb}


SELECT UseCode, CodeDesc, count (CodeDesc)
FROM   parcels LEFT JOIN usecodes
ON     Parcels.UseCode = UseCodes.UseCode2;
GROUP BY UseCode
ORDER BY CodeDesc DESC

```

#2. How many acres are in agricultural preserves?

```{sql connection=parcelsdb}

SELECT sum(Acreage) FROM parcels WHERE AgPres >0


```

#3. What is the mean net assessed value per acre of the entire county?

```{sql connection=parcelsdb}
SELECT sum(Net_AV)/ sum(Acreage) FROM parcels

```


#4. What is the total net assessed value of all non-taxable parcels?

```{sql connection=parcelsdb}
SELECT sum(Net_AV) FROM parcels WHERE NonTaxCode >0
```


#5.What are the 10 largest property holders, by acreage?


```{sql connection=parcelsdb}

SELECT Acreage, M_Address1, M_Address2,sum(Acreage) 
FROM parcels
WHERE M_Address1 AND M_Address2 != 'NA' 
GROUP BY M_Address1
ORDER BY  Acreage DESC
```


#6. What are the 10 largest property holders, by net assessed value?

```{sql connection=parcelsdb}

SELECT Acreage, M_Address1, M_Address2,sum(Net_AV) 
FROM parcels
WHERE M_Address1 AND M_Address2 != 'NA' 
GROUP BY M_Address1
ORDER BY  Net_AV DESC
```
```
